---
title: "PS270 Final Project"
author: "Karis Gilbert"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project Thoughts

I am interested in exploring data related to public opinion on policy issues in the United States. For instance, in a battleground state, what are the voting-eligible population's opinions on contentious issues such as abortion access, medicaid expansion or housing subsidy programs. 

## Research Question + Hypothesis 

Do registered voters' support in Medicaid expansion in their state? More specifically, how do registered voters feel about Medicaid expansion in states that have yet to expand Medicaid? 

As I was looking at a data set from 2021 that was sponsored by he AARP (American Association of Retired Persons), the data is primarily focused on collecting response from registered voters' over the age of 50. Furthermore, the data set is more specialized as I am specifically looking at responses from registered voters over the age of 50 in South Carolina. My hypothesis is that voters over the age of 50 are more inclined to support Medicaid Expansion as they directly benefit from an increase in access to medical programs that are covered by their Medicaid insurance. The two variables I would be most concerned with are the respondent's age and their answer to Q6, from the AARP survey, about the importance of expanding Medicaid for low-income residents in South Carolina A positive result would show a strong correlation between respondents issuing their support for Medicaid expansion and their age as over 50. Additionally, this hypothesis could be further supported by respondents answers to questions about supporting programs for low-income households. If these respondents truly support the expansion of Medicaid they would show support for this program to assist low-income SC residents. On the other hand, my hypothesis would be disproved if despite their age, respondents did not show an overwhelming support for the expansion of Medicaid.

```{r, echo = FALSE}
library(tidyverse)

medicaid <- read_csv("data/medicaid.csv") |>
  rename(Partisanship = d8) |>
  filter(Partisanship != "Don't know or refused [DO NOT READ]",
         Partisanship != "Something else")

```


## Observing Potential Relationship between Partisanship and potential for Medicaid Expansion

1. Assign Numerical Values to Respondent's Answers to Q6: "Do you support or oppose expanding Medicaid health care to South Carolina residents who earn under $18,000 a year, and bringing home hundreds of millions in South Carolina tax dollars from Washington, DC? Is that strongly support, somewhat support, neither support or oppose, somewhat oppose, or strongly oppose?" 

```{r, echo = FALSE}
medicaid <- medicaid |>
  mutate(q6_new = case_when(q6 == "Strongly support" ~ 4,
               q6 == "Somewhat support" ~ 3,
               q6 == "Somewhat oppose" ~ 2,
               q6 == "Strongly oppose" ~ 1,
               q6 == "Don't know or refused [DO NOT READ] Neither support or oppose [DO NOT READ]" ~ NA,
               TRUE ~ NA_real_)) |>
  drop_na(q6_new) |>
  rename(level_of_support = q6_new)

```

```{r, echo = FALSE}
library(knitr)
support_table <- medicaid |> 
  count(level_of_support) |>
  pivot_wider(names_from = level_of_support,
              values_from = n,
              names_prefix = "Support_Level_") 
kable(support_table, caption = "Number of Responses for Each Assigned Level of Support for Medicaid Expansion")
```


## Visualization Observing Respondents Support for Medicaid Expansion Based on Partisanship

```{r fig.width=10, echo = FALSE}
partisanship_plot <- ggplot(data = medicaid,
                        mapping = aes(x = level_of_support, fill = Partisanship)) +
  geom_bar(position = "dodge") + 
  labs(title = "Measuring Support for Medicaid Expansion Against Partisanship",
       x = "Support for Medicaid Expansion in South Carolina 
       (1 = oppose, 2 = somewhat oppose, 3 = somewhat support, 4 = strong support)",
       y = "Number fo Responses")
partisanship_plot
```


## Analysis for Visualization

For the analysis of this variable, support for Medicaid, each level of support was assigned a numerical value in order to designate between opposition and strong support. The scale was used between 1 and 4, with 1 being and the lowest and 4 the highest. Furthermore, the age range of respondents was kept constant as all were over the age of 50. Thus, the two variables being compared were partisanship, given the options of 'Democrat,' 'Independent' or 'Republican, and comparing such to support for expansion of Medicaid in South Carolina to residents that earn less than $18,000 a year. 

As hypothesized, those respondents over the age of 50 demonstrated strong support for Medicaid being expanded to South Carolina residents that earn less than $18,000 a year, partisanship was not a significant contributing variable. Despite 35% of respondents identifying as Republican, 45% strongly supported Medicaid expansion. While those respondents that identified as Democrats reported marginally stronger support over Independent and Republican responses, all three groups demonstrated stronger support than opposition towards Medicaid expansion. Thus to answer the research question, despite differing political ideologies, the strong majority of eligible voters over the age of 50 demonstrate strong support for Medicaid expansion in South Carolina, a state that has not expanded their Medicaid program as of yet. 


## Regression Table 
```{r, echo = FALSE}
library(broom)
library(knitr)

medicaid$Partisanship <- as.factor(medicaid$Partisanship)

model <- lm(level_of_support ~ Partisanship, data = medicaid)

regression_table <- tidy(model)
kable(regression_table, col.names = c("Term", "Estimate", "Std. Error", "t value", "p value"),
      caption = "Regression Results: Support Based on Partisanship")
```

The y-intercept is representative of the average support values for Medicaid expansion as coming from those voters the lean towards democratic partisanship. The reported intercept is 3.82, meaning that on scale measuring for support for Medicaid from 1 to 4, respondents who identify as democratic reported an average score of 3.82 towards supporting policies to expand Medicaid in South Carolina. The score of 3.82 is very close to the maximum score of 4, suggesting that, on average, respondents that identify as Democrats show high support for Medicaid expansion. The coefficient for those respondents identifying as Independents represents how much less this group supports Medicaid expansion compared to the Democratic group. On average, the respondents who aligned as independents reported an average score 0.20 points less than the democratic respondents. Finally, the coefficient for those respondents identifying as Republicans represents how much less this group supports Medicaid expansion compared to the Democratic group. The reported coefficient is -0.35, suggesting that on average, the respondents identifying as Republican support Medicaid expansion 0.35 points less than the Democratic group. 

Moreover, the reported p-values from this particular data set observing the relationship between partisanship and support for Medicaid expansion are significant. The p-value for all three groups, 'Democrats,' 'Independents,' and 'Republicans,' was measured as 0, or 0.0. If we were to use the commonly used significance threshold of 0.05, it can be inferred that a common p-value of 0 across the observed groups indicates highly significant results. The value of 0 is below the alpha level of 0.05, meaning that we can reject the null hypothesis that the effect of the relationship being tested, e.g. partisanship vs. support for Medicaid expansion, is not meaningful. Thus, the observed relationship is highly significant.  


## Another Variable to Consider: Funding

Observing respondent's answers to survey Question 5: "How important is it for state lawmakers to bring South Carolina tax dollars home to expand Medicaid health care? Is it extremely important, very, somewhat, not very, or not at all important?" against Partisanship. 

1. Assign Numerical Values to Respondent's Answers to Q5 
```{r, echo = FALSE}
medicaid <- medicaid |>
  mutate(q5_new = case_when(q5 == "Extremely important" ~ 4,
               q5 == "Very important" ~ 3,
               q5 == "Somewhat important" ~ 2,
               q5 == "Not very or not at all important" ~ 1,
               q5 == "Don't know or refused [DO NOT READ]" ~ NA,
               TRUE ~ NA_real_)) |>
  drop_na(q5_new) |>
  rename(tax_dollars = q5_new)

```

```{r, echo = FALSE}
tax_dollars_table <- medicaid |> 
  count(tax_dollars) |>
  pivot_wider(names_from = tax_dollars,
              values_from = n,
              names_prefix = "Monetary_Support_Level_") 
kable(tax_dollars_table, caption = "Number of Responses for Each Assigned Level of Support 
      for Funding Medicaid from SC Tax Dollars (lowest [1] to highest [5]")

```



2. Visualization Observing Respondents Views towards funding Medicaid based on Partisanship
```{r fig.width=10, echo = FALSE}
funding_plot <- ggplot(data = medicaid,
                        mapping = aes(x = tax_dollars, fill = Partisanship)) +
  geom_bar(position = "dodge") + 
  labs(title = "Measuring Importance of Funding for Medicaid Expansion Against Partisanship",
       x = "Are SC Tax Dollars Necessary for Medicaid Funding 
       (1 = not very important, 2 = somewhat important, 3 = very important, 4 = extremely important)",
       y = "Number fo Responses")
funding_plot
```

Furthermore, in order to consider whether partisanship is truly a significant factor in South Carolina voters' opinions on Medicaid expansion, respondents answers regarding the importance of funding Medicaid was also considered. As with the previous data visualization, responses to question 5, about the importance of funding Medicaid expansion with SC tax dollars, were ranked on a scale from 1 to 4, ranging from 'not very important' to 'extremely important' as answers. These responses were than compared to the same variable, reported partisanship, as in the visual for question 6 from the survey. No respondents viewed the use of funding Medicaid expansion with tax dollars as of no importance, so that assigned value of 1 is absent from the histogram. As with the previous question, respondents identifying as Democrats overwhelmingly view the funding of Medicaid with South Carolina tax dollars as of the most importance. Of the total group of Democratically leaning respondents, around 70% view SC tax dollars as extremely important for Medicaid funding. In comparison, of the total number of respondents identifying as independents, around 48% view SC tax dollars as extremely important for Medicaid funding. For the total number of republican leaning respondents, this percentage was around 39%. Despite differing differing political alignments, the highest percentage of each group supports the use of SC tax dollars as important for an expansion to the state's Medicaid program. 


## Conclusion 

Based on the results of the regression table observing the relationship between partisanship and support for Medicaid, and the relationship between partisanship and funding, it can be suggested that the data strongly supports the hypothesis that respondents over the age of 50 strongly support the expansion of Medicaid in their state. The data set used for analysis kept the age range constant, i.e. all respondents were over the age of 50. Thus, the main variable used for comparison was partisanship. Respondents reported partisanship, given the options of 'Democrat,' 'Independent,' or 'Republican,' was compared to two questions. The first questions observed the relationship between partisanship and whether respondents expressed support for Medicaid expansion in South Carolina to those earning less than $18,000 a year. The next question observed the relationship between partisanship and whether respondents supported the use of tax dollars to fund program expansion in South Carolina. In both questions, despite differing political groups, respondents showed overwhelming support for the expansion of Medicaid in SC to low-income residents and for the program to be funded by tax dollars. Therefore, as hypothesized, despite differences in partisanship registered voters over the age of 50 in South Carolina support the expansion of Medicaid to low-income residents and view the funding of the program from tax dollars as a grave importance in implementing it. 

As my original research question was more broad and looked to observe the responses of registered voters, I would look to expand the data set to include responses from eligible voters under the age of 50. I would still look to separate the results by state and reported partisanship, but it would be interesting to also measure age. Especially given the current climate surrounding voting in the United States, more young people are engaging in voting during the election season as they realize policy concerning the expansion and funding of programs like Medicaid can affect them if they are in a particular socio-economic bracket. As this data set came from AARP, the American Association of Retired Persons, it makes sense that the range of respondents is older, so I would have to look at other organizations that ask similar questions. As I would be changing the age variable, I think I would still want to stick with respondents purely from the state of South Carolina, but possibly another test to consider would be comparing answers towards expanding Medicaid in SC to respondents from other states that still have not expanded their Medicaid provisions. For instance, how do respondents from Wisconsin compare to South Carolina? Do they have similar concerns? Is the partisanship breakdown across respondents vary from state to state? Is that a significant contributing variable? 


